---
title: "A4 BrainAge"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('A4BrainAge.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/A4BrainAge', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr); library(sjPlot)

# Save path
path <- '~/Documents/Projects/AgingBrainStudy/figures/crossectional/'

# Load functions
source('~/Documents/Projects/AgingBrainStudy/scripts/regression_analysis.R')
source('~/Documents/Projects/AgingBrainStudy/scripts/improve_figure.R')

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
theme_set(theme_bw())
```

# BrainAge modelling in A4/Learn study

A linear regression BrainAge model was trained on e4 non-carriers and AB-. The model was then applied to the rest of the cohort.

The model was trained using the following features:

-   Amygdala normalized volume
-   Hippocampus normalized volume
-   Inferior parietal cortical thickness
-   Inferior temporal cortical thickness
-   Middle temporal cortical thickness
-   Entorhinal cortical thickness
-   Parahippocampal cortical thickness
-   Fusiform cortical thickness

Command used: model_age -o analysis/final/a4 -f data/final/a4/processed/structural_features.csv --clinical data/final/a4/processed/clinical.csv

```{r, include = FALSE}
# Load deltas and baseline
deltas <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/final/a4/ageml/model_age/predicted_age.csv') %>%
  rename(delta = delta_all)
baseline <- read.csv('~/Documents/Projects/AgingBrainStudy/data/final/a4/processed/baseline.csv',
                     na.strings = c("", "NA"))

# Merge data
cn_data <- merge(deltas[, c('ID', 'delta')], baseline, by = 'ID')
```

```{r}
# Distribution of deltas for controls
cat("Cognitively Unimpaired \n")
cat("Delta mean:", mean(cn_data$delta), "\n")
cat("Delta std:", sqrt(var(cn_data$delta)), "\n")
```

# Delta differences {.tabset}

## E4 status

```{r}
# Compute group means and their difference
delta_diff_e4 <- abs(diff(aggregate(delta ~ e4_carrier, data = subset(cn_data, !is.na(e4_carrier)), mean)$delta))
p_val_e4 <- t.test(delta ~ e4_carrier, data = cn_data)$p.value

# Plot
ggplot(subset(cn_data, !is.na(e4_carrier)), aes(x = e4_carrier, y = delta)) +
  geom_boxplot() +
  labs(
    title = "Delta differences by E4 status",
    x = "E4 status",
    y = "Delta"
  ) +
  annotate(
    "text",
    x = 1.5, 
    y = 6,
    label = sprintf("Mean difference: %.3f", delta_diff_e4),
    size = 4,
    fontface = "italic"
  ) +
  annotate(
    "text",
    x = 1.5, 
    y = 5,
    label = sprintf("p = %.2g", p_val_e4),
    size = 4,
    fontface = "italic"
  )
```

## Amyloid status

```{r}
# Compute group means and their difference
delta_diff_ab <- abs(diff(aggregate(delta ~ ab_status, data = subset(cn_data, !is.na(ab_status)), mean)$delta))
p_val_ab <- t.test(delta ~ ab_status, data = cn_data)$p.value

# Plot
ggplot(subset(cn_data, !is.na(ab_status)), aes(x = ab_status, y = delta)) +
  geom_boxplot() +
  labs(
    title = "Delta differences by AB status",
    x = "AB status",
    y = "Delta"
  ) +
  annotate(
    "text",
    x = 1.5, y = 6,  # Position the text at the top center
    label = sprintf("Mean difference: %.3f", delta_diff_ab),
    size = 4,
    fontface = "italic"
  ) +
  annotate(
    "text",
    x = 1.5, y = 5,
    label = sprintf("p = %.2g", p_val_ab),
    size = 4,
    fontface = "italic"
  )
```

## Sex differences

```{r}
# Compute group means and their difference
delta_diff_sex <- abs(diff(aggregate(delta ~ sex, data = subset(cn_data, !is.na(sex)), mean)$delta))
p_val_sex <- t.test(delta ~ sex, data = cn_data)$p.value

# Plot
ggplot(subset(cn_data, !is.na(sex)), aes(x = sex, y = delta)) +
  geom_boxplot() +
  labs(
    title = "Delta differences by Sex",
    x = "Sex",
    y = "Delta"
  ) +
  annotate(
    "text",
    x = 1.5, y = 6,  # Position the text at the top center
    label = sprintf("Mean difference: %.3f", delta_diff_sex),
    size = 4,
    fontface = "italic"
  ) +
  annotate(
    "text",
    x = 1.5, y = 5,
    label = sprintf("p = %.2g", p_val_sex),
    size = 4,
    fontface = "italic"
  )
```

# Biomarkers {.tabset}

Correlation between BrainAge delt and different biomarkers. Sensitivity analysis includes addin sex, educaiton, APOEe4, AB-status and MRI Age as covariates. Further sensitivity analysis limits datapoints to those within 1 year of the MRI.

## PACC

```{r}
# Run analysis on PACC
pacc_analysis = regression_analysis(cn_data, 
                                    'PACC_mri', 
                                    'delta', 
                                    c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Plot results
pacc_analysis$scatter_plot

# Time sensitivity analysis
pacc_sens = regression_analysis(cn_data[cn_data$time_diff_pacc <=1 & cn_data$time_diff_pacc >=-1,],
                                'PACC_mri', 
                                'delta', 
                                c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Sensitivity analysis
cat("Sensitivity analysis: \n")
cat("p-value of delta including covariates:", 
    pacc_analysis$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")
cat("p-value of delta including covariates and time difference:", 
    pacc_sens$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")

# Apply figure improvement and save
pacc_analysis$scatter_plot <- improve_graph(pacc_analysis$scatter_plot, 'Delta (years)', 'PACC')
ggsave(paste(path, 'pacc_a4.png'), pacc_analysis$scatter_plot, 
       width = 50, height = 36, units = "mm", 
       dpi = 500,     # Higher resolution
       scale = 1) 
```

## AB

```{r}
# Run analysis on AB
ab_analysis = regression_analysis(cn_data[cn_data$secondary == 1,],
                                  'ab_composite', 
                                  'delta', 
                                  c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Plot results
ab_analysis$scatter_plot

# Time sensitivity analysis
ab_sens = regression_analysis(cn_data[cn_data$secondary == 1 & cn_data$time_diff_ab <=1 & cn_data$time_diff_ab >=-1,],
                              'ab_composite', 
                              'delta', 
                              c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Sensitivity analysis
cat("Sensitivity analysis: \n")
cat("p-value of delta including covariates:", 
    ab_analysis$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")
cat("p-value of delta including covariates and time difference:", 
    ab_sens$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")

# Apply figure improvement and save
ab_analysis$scatter_plot <- improve_graph(ab_analysis$scatter_plot, 'Delta (years)', 'Aβ-PET')
ggsave(paste(path, 'ab_a4.png'), ab_analysis$scatter_plot, 
       width = 50, height = 36, units = "mm", 
       dpi = 500,     # Higher resolution
       scale = 1) 
```

## ptau217

```{r}
# Run analysis on AB
ptau_analysis = regression_analysis(cn_data[cn_data$exploratory == 1,],
                                  'ptau', 
                                  'delta', 
                                  c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Plot results
ptau_analysis$scatter_plot

# Time sensitivity analysis
ptau_sens = regression_analysis(cn_data[cn_data$exploratory == 1 & cn_data$time_diff_ptau <=1 & cn_data$time_diff_ptau >=-1,],
                              'ptau', 
                              'delta', 
                              c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Sensitivity analysis
cat("Sensitivity analysis: \n")
cat("p-value of delta including covariates:", 
    ptau_analysis$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")
cat("p-value of delta including covariates and time difference:", 
    ptau_sens$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")

# Apply figure improvement and save
ptau_analysis$scatter_plot <- improve_graph(ptau_analysis$scatter_plot, 'Delta (years)', 'pTau217')
ggsave(paste(path, 'ptau_a4.png'), ptau_analysis$scatter_plot, 
       width = 50, height = 36, units = "mm", 
       dpi = 500,     # Higher resolution
       scale = 1) 
```

## TAU

```{r}
# Run analysis on AB
tau_analysis = regression_analysis(cn_data[cn_data$exploratory == 1,],
                                  'tau_composite', 
                                  'delta', 
                                  c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Plot results
tau_analysis$scatter_plot

# Time sensitivity analysis
tau_sens = regression_analysis(cn_data[cn_data$exploratory == 1 & cn_data$time_diff_tau <=1 & cn_data$time_diff_tau >=-1,],
                              'tau_composite', 
                              'delta', 
                              c('sex', 'edu', 'e4_carrier', 'ab_status', 'mri_age'))

# Sensitivity analysis
cat("Sensitivity analysis: \n")
cat("p-value of delta including covariates:", 
    tau_analysis$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")
cat("p-value of delta including covariates and time difference:", 
    tau_sens$model_full_summary$coefficients["delta", "Pr(>|t|)"], "\n")

# Apply figure improvement and save
tau_analysis$scatter_plot <- improve_graph(tau_analysis$scatter_plot, 'Delta (years)', 'Tau-PET')
ggsave(paste(path, 'tau_a4.png'), tau_analysis$scatter_plot, 
       width = 50, height = 36, units = "mm", 
       dpi = 500,     # Higher resolution
       scale = 1) 
```