---
title: "Clinical Enrichment with BrainAge"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    self_contained: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('ClinicalEnrichment.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/ClinicalEnrichment', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr);library(nlme); library(sjPlot);library(longpower);library(ggeffects);library(emmeans);library(gridExtra)

# Set global options
emm_options(lmerTest.limit = 15872)
emm_options(pbkrtest.limit = 15872)

# Load cost of clinical trial
source('~/Documents/Projects/AgingBrainStudy/scripts/calculate_drug_trial_costs.R')

# Path to save figures
path = '~/Documents/Projects/AgingBrainStudy/figures/clinical_enrichment/'

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
theme_set(theme_bw())

graph_th <- theme(
      text = element_text(family = "Times New Roman", size = 12),
      axis.text = element_text(family = "Times New Roman", size = 12),
      axis.title = element_text(family = "Times New Roman", size = 12),
      strip.text = element_text(family = "Times New Roman", size = 8),
      plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
      axis.title.y = element_text(margin = margin(r = 2, unit = "mm")),
      axis.title.x = element_text(margin = margin(t = 2, unit = "mm"))
    ) 
```

# Brain Age in Clinical Trials

We will explore usign BrainAge as a clinical enrichment biomarker in two clinical trials:

-   ADNI: MCI participants with amyloid positivity in two year trial looking at CDR sum of boxes
-   A4: Preclinical AD participants with amyloid positivity in five year trial looking at PACC

```{r}
# Trial variables
treatment_eff = 0.27
power = 0.9
sig_level = 0.05
n = NULL
method = 'edland'

# Lecanemab costs
screened_lecanemab = 5967
recruited_lecanemab = 1796
treatment_cost_per_year = 26500
scren_cost = 6957
screen_cost_mri = 353
screen_pass_lecanemab = recruited_lecanemab / screened_lecanemab
```

# ADNI

```{r}

# Load ADNI data for MCI
data_adni <- read.csv('~/Documents/Projects/AgingBrainStudy/data/final/adni/processed/longitudinal.csv')

# Filter only MCI, Amyloid Positive and first 24 months
data <- data_adni %>% 
  filter(VISCODE %in% c('bl', 'm06', 'm12', 'm18', 'm24')) %>%
  filter(ab_status == 'ab+') %>%
  filter(diagnosis == 'MCI')

# Extract those with delta_all above median
data_median <- data %>% 
  filter(delta > median(data$delta))

# Total number of unique participants in each
n_adni_participants <- length(unique(data$ID))
n_adni_participants_median <- length(unique(data_median$ID))

# Outcome of interest
outcome = 'CDRSUM'
```

## Linear Mixed Effects Model {.tabset}

Simple linear mixed effects model with random intercept and slope for each particpant for the three subsets.

### All

```{r}
# Create model
lmm_adni_all <-lmer(as.formula(paste(outcome, "~ time_mri + (1 + time_mri | ID)")), data = data)

# Plot model
adni_all_p <- plot(ggpredict(lmm_adni_all, terms = c("time_mri [all]"))) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, max(data$time_mri))) + 
  scale_y_continuous(limits = c(0, 4.5)) +
  labs(
    title = "",
    y = "CDRSUM",
    x = "Time from MRI (years)",
  )
  
print(adni_all_p)

# Save model
adni_all_p <- adni_all_p + graph_th
ggsave(paste(path, 'lme_adni_all.png'), adni_all_p, 
       width = 70, height = 70, units = "mm", 
       dpi = 500,   # Higher resolution
       scale = 1)
```

### Median

```{r}
# Create model
lmm_adni_median <-lmer(as.formula(paste(outcome, "~ time_mri + (1 + time_mri | ID)")), data = data_median)

# Plot model
adni_median_p <- plot(ggpredict(lmm_adni_median, terms = c("time_mri [all]"))) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, max(data$time_mri))) + 
  scale_y_continuous(limits = c(0, 4.5)) +
  labs(
    title = "",
    y = "CDRSUM",
    x = "Time from MRI (years)",
  )
  
print(adni_median_p)

# Save model
adni_median_p <- adni_median_p + graph_th
ggsave(paste(path, 'lme_adni_median.png'), adni_median_p, 
       width = 70, height = 70, units = "mm", 
       dpi = 500,   # Higher resolution
       scale = 1)
```

## Sample and power caclulations {.tabset}

The trial will last 2 years with follow ups every 6 months.

```{r}
# Set parameters for ADNI two years
t_visits_adni = seq(0, 2, 0.5)
```

### All

```{r}
# calculate power / sample size using longpower package
pwr_all <- lmmpower(lmm_adni_all, pct.change = treatment_eff, t = t_visits_adni, power = power, sig.level = sig_level, n = n, method = method)

# Print number of subjects
n_adni_all <- pwr_all$N
cat("Number of participants required for clinical trial: ", n_adni_all, "\n")
```

### Median

```{r}
# calculate power / sample size using longpower package
pwr_median <- lmmpower(lmm_adni_median, pct.change = treatment_eff, t = t_visits_adni, power = power, sig.level = sig_level, n = n, method = method)

# Print number of subjects
n_adni_median <- pwr_median$N
cat("Number of participants required for clinical trial: ", n_adni_median, "\n")

# Reduction percentage
reduction_median <- (n_adni_all - n_adni_median) / n_adni_all * 100
cat("Reduction in number of participants required: ", reduction_median, "%\n")
```

## Calculate savings {.tabset}

Calculate costs of the trail

### All

```{r}

#Calculate cost
cost_adni_all <- calculate_drug_trial_costs(
  years = 2,
  participants_required = n_adni_all,
  screen_pass_rate = screen_pass_lecanemab,
  screen_cost = scren_cost,
  mri_cost = screen_cost_mri,
  treatment_cost_per_year = treatment_cost_per_year,
  has_mri_screening = FALSE,
  mri_pass_rate = 0
)

cat("Total cost of the trial: ", cost_adni_all$total_cost, "\n")
cat("Screening cost: ", cost_adni_all$screening_cost, "\n")
cat("Participants to screen: ", cost_adni_all$participants_to_screen, "\n")
cat("Treatment cost: ", cost_adni_all$treatment_cost, "\n")
```

### Median

```{r}

# Calculate pass rate for MRI
mri_pass_rate_median = n_adni_participants_median / n_adni_participants
cat("MRI pass rate: ", mri_pass_rate_median, "\n")

# Calculate cost
cost_adni_median <- calculate_drug_trial_costs(
  years = 2,
  participants_required = n_adni_median,
  screen_pass_rate = screen_pass_lecanemab,
  screen_cost = scren_cost,
  mri_cost = screen_cost_mri,
  treatment_cost_per_year = treatment_cost_per_year,
  has_mri_screening = TRUE,
  mri_pass_rate = mri_pass_rate_median
)

cat("Total cost of the trial: ", cost_adni_median$total_cost, "\n")
cat("Screening cost: ", cost_adni_median$screening_cost, "\n")
cat("Participants to screen: ", cost_adni_median$participants_to_screen, "\n")
cat("Treatment cost: ", cost_adni_median$treatment_cost, "\n")
```


# A4

Looking at AB+ with placebo treatment in the A4 trial. We will look at the PACC outcome.

```{r}
# Load A4 data
data_a4 <- read.csv('~/Documents/Projects/AgingBrainStudy/data/final/a4/processed/longitudinal.csv')

# Placebo data
data_placebo <- data_a4 %>% 
  filter(cohort == 'A4 Placebo') %>% # Only placebo
  filter(!is.na(PACC)) %>%
  filter(VISCODE <= 60) # Only before the first 4 years

# Extract those with delta_all above median
data_a4_median <- data_placebo %>% 
  filter(delta > median(data_placebo$delta, na.rm = TRUE))

# Total number of unique participants in each
n_a4_participants <- length(unique(data_placebo$ID))
n_a4_participants_median <- length(unique(data_a4_median$ID))

# Outcome of interest
outcome = 'PACC'
```

## Linear Mixed Effects Model {.tabset}

Simple linear mixed effects model with random intercept and slope for each particpant for the three subsets.

### All

```{r}
# Create model

lmm_a4_all <-lmer(as.formula(paste(outcome, "~ I(time_mri^2) + time_mri + LMSTORY + Cumulative_Dose_Scaled + (1 + time_mri | ID)")), data = data_placebo)

# Plot model
a4_all_p <- plot(ggpredict(lmm_a4_all, terms = c("time_mri [all]"))) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, max(data_placebo$time_mri))) + 
  scale_y_continuous(limits = c(-1.5, 1)) +
  labs(
    title = "",
    y = "PACC",
    x = "Time from MRI (years)",
  )
  
print(a4_all_p)

# Save model
a4_all_p <- a4_all_p + graph_th
ggsave(paste(path, 'lme_a4_all.png'), a4_all_p, 
       width = 70, height = 70, units = "mm", 
       dpi = 500,   # Higher resolution
       scale = 1)
```

### Median

```{r}
# Create model
lmm_a4_median <-lmer(as.formula(paste(outcome, "~ I(time_mri^2) + time_mri + LMSTORY + Cumulative_Dose_Scaled + (1 + time_mri | ID)")), data = data_a4_median)

# Plot model
a4_median_p <- plot(ggpredict(lmm_a4_median, terms = c("time_mri [all]"))) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, max(data_placebo$time_mri))) + 
  scale_y_continuous(limits = c(-1.5, 1)) +
  labs(
    title = "",
    y = "PACC",
    x = "Time from MRI (years)",
  )
  
print(a4_median_p)

# Save model
a4_median_p <- a4_median_p + graph_th
ggsave(paste(path, 'lme_a4_median.png'), a4_median_p, 
       width = 70, height = 70, units = "mm", 
       dpi = 500,   # Higher resolution
       scale = 1)
```

## Sample and power caclulations {.tabset}

Trial duration is 4 years with follow ups every 6 months.

```{r}
# Set parameters for A4 5 years
trial_time = 4 # years
n_visit_per_year = 2 
t_visits_a4 = seq(0, trial_time, 1/n_visit_per_year)
```

### All

```{r}
# First calculate linear slope of change from baseline to trial end
pacc_baseline_all<- fixef(lmm_a4_median)["(Intercept)"]
pacc_end_trial_all <- summary(emmeans(lmm_a4_all, specs = ~ time_mri, at = list(time_mri = trial_time, pbkrtest.limit = 15872)))$emmean
delta_change_per_year_all <- (pacc_end_trial_all - pacc_baseline_all) / trial_time

# Calculate treatment effect
effect_all = delta_change_per_year_all * treatment_eff

# Power calculation
pwr_all_a4 <- lmmpower(lmm_a4_all, delta = effect_all, t = t_visits_a4, power = power, sig.level = sig_level, n = n, method = method)

# Print number of subjects
n_a4_all <- pwr_all_a4$N
cat("Number of participants required for clinical trial: ", n_a4_all, "\n")
```

### Median

```{r}
# First calculate linear slope of change from baseline to trial end
pacc_baseline_median <- fixef(lmm_a4_median)["(Intercept)"]
pacc_end_trial_median <- summary(emmeans(lmm_a4_median, specs = ~ time_mri, at = list(time_mri = trial_time, pbkrtest.limit = 15872)))$emmean
delta_change_per_year_median <- (pacc_end_trial_median - pacc_baseline_median) / trial_time

# Calculate treatment effect
effect_median = delta_change_per_year_median * treatment_eff

# Power calculation
pwr_median_a4 <- lmmpower(lmm_a4_median, delta = effect_median, t = t_visits_a4, power = power, sig.level = sig_level, n = n, method = method)

# Print number of subjects
n_a4_median <- pwr_median_a4$N
cat("Number of participants required for clinical trial: ", n_a4_median, "\n")

# Reduction percentage
reduction_median <- (n_a4_all - n_a4_median) / n_a4_all * 100
cat("Reduction in number of participants required: ", reduction_median, "%\n")

# Screen fails for MRI
mri_pass_rate_median = n_a4_participants_median / n_a4_participants
cat("MRI pass rate: ", mri_pass_rate_median, "\n")
```

# Ptau217 screening

```{r}
# Obtain subjects with ptau217 data
data_a4_ptau217 <- data_a4 %>% 
  filter(!is.na(ptau)) %>% # Have ptau217
  filter(VISCODE <= 60) # Only before the first 5 years
n_a4_ptau217_participants <- length(unique(data_a4_ptau217$ID))

# Obtain baseline data by grouping by ID, ordering by viscode and selecting firtst row
data_a4_ptau217_baseline <- data_a4_ptau217 %>% 
  group_by(ID) %>% 
  arrange(VISCODE) %>% 
  slice(1) %>% 
  ungroup()

# Find quantiles for pTau217
ptau217_quantiles <- quantile(data_a4_ptau217_baseline$ptau, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
delta_quantiles <- quantile(data_a4_ptau217_baseline$delta, probs = c(0.0, 0.25, 0.5, 0.75), na.rm = TRUE)
```

```{r}
# Trial variables
treatment_eff = 0.30
power = 0.8
sig_level = 0.05
n = NULL
method = 'edland'
outcome = 'PACC'
trial_time = 4 # years
n_visit_per_year = 2 
t_visits_a4 = seq(0, trial_time, 1/n_visit_per_year)

# Standard trail
lme_ptau_all <-lmer(as.formula(paste(outcome, "~ I(time_mri^2) + time_mri + LMSTORY + Cumulative_Dose_Scaled + time_mri*cohort +  (1 + time_mri | ID)")), data = data_a4_ptau217)

# Calculate baseline and end trial values
pacc_baseline_ptau_all <- fixef(lme_ptau_all)["(Intercept)"]
pacc_end_trial_ptau_all <- summary(emmeans(lme_ptau_all, specs = ~ time_mri, at = list(time_mri = trial_time, pbkrtest.limit = 15872)))$emmean
delta_change_per_year_ptau_all <- (pacc_end_trial_ptau_all - pacc_baseline_ptau_all) / trial_time

# Calculate treatment effect
effect_ptau_all = delta_change_per_year_ptau_all * treatment_eff

# Calculate power / sample size using longpower package
pwr_ptau_all <- lmmpower(lme_ptau_all, delta = effect_ptau_all, t = t_visits_a4, power = power, sig.level = sig_level, n = n, method = method)
n_a4_ptau_all <- pwr_ptau_all$N
```

```{r}
# Initialize results storage
results_ptau_df <- data.frame(
  ptau_threshold = numeric(),
  delta_threshold = numeric(),
  n_participants = integer(),
  n_trial_required = numeric(),
  reduction_percentage = numeric(),
  stringsAsFactors = FALSE
)

# Print header
cat("\n", paste(rep("=", 80), collapse=""), "\n")
cat("POWER ANALYSIS RESULTS\n")
cat(paste(rep("=", 80), collapse=""), "\n")
cat(sprintf("%-12s %-12s %-12s %-15s %-12s %-12s %-12s %-15s %-15s \n", 
            "pTau217", "Delta", "Available", "Trial Required", "Reduction", "CI Lower", "CI Upper", "Reduction Lower", "Reduction Upper"))
cat(sprintf("%-12s %-12s %-12s %-15s %-12s %-12s %-12s %-15s %-15s \n", 
            "Threshold", "Threshold", "Participants", "Sample Size", "(%)", "Sample Size", "Sample Size", "(%)", "(%)"))
cat(paste(rep("-", 80), collapse=""), "\n")

# Do for loop over ptau217 quantiles and delta quantiles
for (ptau_quantile in ptau217_quantiles) {
  for (delta_quantile in delta_quantiles) {
    # Filter data
    data_filtered <- data_a4_ptau217 %>%
      filter(ptau >= ptau_quantile, delta >= delta_quantile)
    
    # Model
    lme_ptau <- lmer(as.formula(paste(outcome, "~  I(time_mri^2) + time_mri + LMSTORY + Cumulative_Dose_Scaled + time_mri*cohort + LMSTORY + (1 + time_mri | ID)")), data = data_filtered)
    
    # Calculate baseline and end trial values
    pacc_baseline_ptau <- fixef(lme_ptau)["(Intercept)"]
    pacc_end_trial_ptau <- summary(emmeans(lme_ptau, specs = ~ time_mri, at = list(time_mri = trial_time, pbkrtest.limit = 15872)))$emmean
    delta_change_per_year_ptau <- (pacc_end_trial_ptau - pacc_baseline_ptau) / trial_time
    
    # Calculate treatment effect
    effect_ptau = delta_change_per_year_ptau * treatment_eff
    
    # Calculate power / sample size
    pwr_ptau <- lmmpower(lme_ptau, delta = effect_ptau, t = t_visits_a4, power = power, sig.level = sig_level, n = n, method = method)
    n_a4_ptau <- pwr_ptau$N
    n_percentage <- n_a4_ptau / n_a4_ptau_all * 100
    
    # Calculate confidence intervals
    ni_ci_upper <- 2*pwr_ptau$n.CI[1]
    ni_ci_lower <- 2*pwr_ptau$n.CI[3]
    reduction_lower <-  ni_ci_lower / n_a4_ptau_all * 100
    reduction_upper <-  ni_ci_upper / n_a4_ptau_all * 100
    
    # Calculate number of participants
    n_participants <- length(unique(data_filtered$ID))
    
    # Store results
    results_ptau_df <- rbind(results_ptau_df, data.frame(
      ptau_threshold = ptau_quantile,
      delta_threshold = delta_quantile,
      n_participants = n_participants,
      n_trial_required = n_a4_ptau,
      reduction_percentage = n_percentage,
      n_ci_lower = ni_ci_lower,
      n_ci_upper = ni_ci_upper,
      reduction_lower = reduction_lower,
      reduction_upper = reduction_upper
    ))
    
    # Print results in formatted table
    cat(sprintf("%-12.2f %-12.2f %-12d %-15.0f %-12.1f %-15.0f %-15.0f %-12.1f %-12.1f \n ", 
                ptau_quantile, delta_quantile, n_participants, n_a4_ptau, n_percentage, ni_ci_lower, ni_ci_upper, reduction_lower, reduction_upper))
  }
}

# Optional: Show summary statistics
cat("SUMMARY STATISTICS:\n")
cat(sprintf("Trial sample size range: %.0f - %.0f participants\n", 
            min(results_ptau_df$n_trial_required), max(results_ptau_df$n_trial_required)))
cat(sprintf("Available participants range: %d - %d\n", 
            min(results_ptau_df$n_participants), max(results_ptau_df$n_participants)))
cat(sprintf("Sample size reduction range: %.1f%% - %.1f%%\n", 
            min(results_ptau_df$reduction_percentage), max(results_ptau_df$reduction_percentage)))
```


```{r}
graph_th_screening <- theme(
      title = element_text(family = "Times New Roman", size = 10),
      plot.title = element_text(hjust = 0.5),
      text = element_text(family = "Times New Roman", size = 10),
      axis.text = element_text(family = "Times New Roman", size = 10),
      axis.title = element_text(family = "Times New Roman", size = 10),
      strip.text = element_text(family = "Times New Roman", size = 8),
      plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
      axis.title.y = element_text(margin = margin(r = 2, unit = "mm")),
      axis.title.x = element_text(margin = margin(t = 2, unit = "mm"))
    ) 


ct_plot <- function(panel_data, title) {
  
  # Create plot data using existing reduction_percentage values
  # We need 5 points: All (100%), then the 4 combinations from panel_data
  plot_data <- data.frame(
    step_numeric = c(1, 2, 3, 3, 3),  # All, Plasma, 3x BrainAge
    participants_pct = c(100, panel_data$reduction_percentage),  # Start with 100% for "All"
    participants_ci_low = c(100, panel_data$reduction_lower),
    participants_ci_high = c(100, panel_data$reduction_upper)
  )
  
  # Create separate data for lines
  red_line_data <- data.frame(x = c(1, 2), y = c(100, plot_data$participants_pct[2]))
  green_line1_data <- data.frame(x = c(2, 3), y = c(plot_data$participants_pct[2], plot_data$participants_pct[3]))
  green_line2_data <- data.frame(x = c(2, 3), y = c(plot_data$participants_pct[2], plot_data$participants_pct[4]))  
  green_line3_data <- data.frame(x = c(2, 3), y = c(plot_data$participants_pct[2], plot_data$participants_pct[5]))
  point_data <- data.frame(
    x = c(1, 2, 3, 3, 3),
    y = c(100, plot_data$participants_pct[2:5]),
    y_min = c(100, plot_data$participants_ci_low[2:5]),
    y_max = c(100, plot_data$participants_ci_high[2:5]),
    color = c("grey50", "red", "darkgreen", "darkgreen", "darkgreen")
  )
  
  # Create the plot
  p <- ggplot() +
    # Red line for plasma step
    geom_line(aes(x = x, y = y), data = red_line_data, color = "red", size = 1.2) +
    
    # Green lines for BrainAge steps (different line types)
    geom_line(aes(x = x, y = y), data = green_line1_data, color = "darkgreen", size = 1.2, linetype = "solid") +
    geom_line(aes(x = x, y = y), data = green_line2_data, color = "darkgreen", size = 1.2, linetype = "dashed") +
    geom_line(aes(x = x, y = y), data = green_line3_data, color = "darkgreen", size = 1.2, linetype = "dotted") +
    
    # Points
    geom_point(aes(x = x, y = y, color = color), data = point_data, size = 3) +
    geom_errorbar(aes(x = x, ymin = y_min, ymax = y_max, color = color), 
                  data = point_data, width = 0.1, size = 0.5) +
    scale_color_identity() +
    
    # Percentage labels
    geom_text(aes(x = 2.3, y = plot_data$participants_pct[2], 
                  label = paste0(round(abs(plot_data$participants_pct[2]), 0), "%")), 
              vjust = -2, size = 2, color = "red") +
    geom_text(aes(x = rep(3.35, 3), y = plot_data$participants_pct[3:5], 
                  label = paste0(round(abs(plot_data$participants_pct[3:5]), 0), "%")), 
                 size = 2, color = "darkgreen") +
    geom_text(aes(x = 1, y = 100, label = "100%"), 
              vjust = -2, size = 2, color = "gray50") +
    
    # Horizontal reference line at 100
    geom_hline(yintercept = 100, linetype = "dashed", color = "gray50", alpha = 0.7) +
    
    # Styling
    scale_x_continuous(breaks = c(1, 2, 3), labels = c("All", "Plasma", "BrainAge"), limits = c(0.8, 3.5)) +
    scale_y_continuous(limits = c(0, 105), breaks = seq(0, 100, 25)) +
    labs(title = title, 
         x = "Step", 
         y = "Participants (%)") +
    theme_minimal() + graph_th_screening
  
  return(p)
}

```

```{r}
# Example usage:
# For Q2-Q4 panel
ptau_threshold_q2q4 <- ptau217_quantiles[1]
panel_data_q2q4 <- results_ptau_df %>% filter(ptau_threshold == ptau_threshold_q2q4)
p1 <- ct_plot(panel_data_q2q4, "Plasma p-tau217\nQ2-Q4")

# For Q3-Q4 panel
ptau_threshold_q3q4 <- ptau217_quantiles[2]
panel_data_q3q4 <- results_ptau_df %>% filter(ptau_threshold == ptau_threshold_q3q4)
p2 <- ct_plot(panel_data_q3q4, "Plasma p-tau217\nQ3-Q4")

# For Q4 panel
ptau_threshold_q4 <- ptau217_quantiles[3]
panel_data_q4 <- results_ptau_df %>% filter(ptau_threshold == ptau_threshold_q4)
p3 <- ct_plot(panel_data_q4, "Plasma p-tau217\nQ4")

# Combine all plots into one figure
final_plot <- arrangeGrob(p1, p2, p3, ncol = 3)

# Add legend manually (since we have custom colors and lines)
legend_plot <- ggplot() +
  geom_line(aes(x = c(1,2), y = c(1,1)), color = "red", size = 1.2) +
  geom_line(aes(x = c(1,2), y = c(0.5,0.5)), color = "darkgreen", size = 1.2) +
  geom_text(aes(x = 2.2, y = 1), label = "Step 1: plasma p-tau217", size = 3, hjust = 0) +
  geom_text(aes(x = 2.2, y = 0.5), label = "Step 2: BrainAge delta", size = 3, hjust = 0) +
  geom_text(aes(x = 4, y = 0.5), label = "BrainAge delta quartiles — Q2-Q4   - - Q3-Q4   ··· Q4", size = 3, hjust = 0) +
  xlim(0, 8) + ylim(0, 1.5) +
  theme_void()

# Display the final plot with legend
grid.arrange(final_plot, legend_plot, heights = c(4, 0.5))

# Save the final plot
ggsave(paste(path, 'ptau217_screening.png'), 
       arrangeGrob(final_plot, legend_plot, heights = c(4, 0.5)), 
       width = 160, height = 100, units = "mm", 
       dpi = 500,   # Higher resolution
       scale = 1)
```