---
title: "HABSBrainAge"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('HABSBrainAge.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/HABSBrainAge', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr); library(sjPlot)


knitr::opts_chunk$set(echo = FALSE, message = FALSE)
theme_set(theme_bw())
```

# BrainAge modelling in HABS

A linear regression BrainAge model was trained on e4 non-carriers and AB-. The model was then applied to the rest of the cohort.

The model was trained using the following features:

-   Amygdala normalized volume
-   Hippocampus normalized volume
-   Inferior parietal cortical thickness
-   Inferior temporal cortical thickness
-   Middle temporal cortical thickness
-   Entorhinal cortical thickness
-   Parahippocampal cortical thickness
-   Fusiform cortical thickness

Command used: model_age -o analysis/HABS/structural_brainage -f data/HABS/processed/structural_features.csv --clinical data/HABS/processed/clinical.csv

```{r, include = FALSE}
# Load deltas
deltas <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/HABS/structural_brainage/ageml/model_age/predicted_age.csv')
deltas <- deltas %>% 
  rename(delta = delta_all, predicted_age = predicted_age_all)

# Obtain deltas trained with A4
deltas_trained_wA4 <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/A4_HABS_joint/structural_brainage/ageml/model_age/predicted_age.csv')
deltas_trained_wA4 <- deltas_trained_wA4[deltas_trained_wA4$SubjIDshort %in% grep('_', deltas_trained_wA4$SubjIDshort, value = TRUE),]
deltas_trained_wA4 <- deltas_trained_wA4 %>% 
  rename(predicted_age_trained_wA4 = predicted_age_all, delta_wA4 = delta_all) %>%
  select(SubjIDshort, predicted_age_trained_wA4, delta_wA4)
data <- merge(deltas, deltas_trained_wA4, by = 'SubjIDshort', all = TRUE)

# Load factors
factors <- read.csv('~/Documents/Projects/AgingBrainStudy/data/HABS/processed/factors.csv', na.strings = c("", " ", "NaN", "NA", NA))
data <- merge(data, factors, by = 'SubjIDshort', all = TRUE)
```

```{r}
# Distribution of deltas
mean <- mean(data$delta, na.rm=TRUE)
print(mean)
std <- sqrt(var(data$delta, na.rm=TRUE))
print(std)
```

## Comparison of BrainAge models

We also trained a model on A4 and applied it to HABS.

Command used: model_age -o analysis/A4_HABS_joint/structural_brainage -f data/A4_HABS_joint/processed/structural_features.csv --clinical data/A4_HABS_joint/A4_as_cn.csv

```{r}
# Plot deltas vs deltas_trained_wA4
ggplot(data = data, aes(x = delta, y = delta_wA4)) +
  geom_point() +
  stat_cor() +
  labs(x = 'Delta', y = 'Delta A4 trained', title = 'Comparison of BrainAge models')

# histogram of deltas coloured by model
ggplot(data = data, aes(x = delta, fill = 'delta')) +
  geom_histogram(alpha = 0.5, position = 'identity') +
  geom_histogram(data = data, aes(x = delta_wA4, fill = 'delta_wA4'), alpha = 0.5, position = 'identity') +
  labs(x = 'Delta', y = 'Count', title = 'Histogram of deltas coloured by model') +
  scale_fill_manual(values = c('blue', 'red'))
```

```{r}
# Caculate mean absolute error for each model
mae <- data %>% 
  summarise(mae = mean(abs(age- predicted_age), na.rm = TRUE),
            mae_trained_wA4 = mean(abs(age - predicted_age_trained_wA4), na.rm = TRUE))
mae
std <- data %>% 
  summarise(std = sd(delta, na.rm = TRUE),
            std_trained_wA4 = sd(delta_wA4, na.rm = TRUE))
std
```

```{r}
regression_analysis <- function(data, x, y, covars) {
  
  # Filter nan values
  data <- data[complete.cases(data[, c(x, y, covars)]), ]
  
  # Scatter plot with correlation
  scatter_plot <- ggplot(data, aes(x = !!sym(y), y = !!sym(x))) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
    
  # Create a formula for the regression model
  covar_formula <- paste(covars, collapse = " + ")
  full_formula <- as.formula(paste(x, "~", y, "+", covar_formula))
  
  # Linear regression model
  ols_results <- lm(as.formula(paste(x, "~", y)), data = data)
  ols_full_results <- lm(full_formula, data = data)
  
  # Model summary
  model_summary <- summary(ols_results)
  model_full_summary <- summary(ols_full_results)
  
  # Plot model
  pred_plot <- plot_model(ols_results, type = c('pred'), terms = c(y), show.values = TRUE) +
    geom_point(data = data, aes(x = !!sym(y), y = !!sym(x)), 
             inherit.aes = F) + 
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
  
  # Return a list with all results
  return(list(
    scatter_plot = scatter_plot,
    simple_model = ols_results,
    regression_model = ols_full_results,
    model_summary = model_summary,
    model_full_summary = model_full_summary,
    predicted_values_plot = pred_plot
  ))
}
```

## Delta differences {.tabset}

### E4 status

```{r}
ggplot(subset(data, !is.na(E4_Status)), aes(x = E4_Status, y = delta)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by E4 status',
       x = 'E4 status',
       y = 'Delta')
```

### Amyloid status

```{r}
ggplot(subset(data, !is.na(PIB_FS_SUVR_Group)), aes(x = PIB_FS_SUVR_Group, y = delta)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by Amyloid status',
       x = 'Amyloid status',
       y = 'Delta')
```

### Sex differences

```{r}
ggplot(subset(data, !is.na(SEX)), aes(x = SEX, y = delta)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by Amyloid status',
       x = 'SEX',
       y = 'Delta')
```

## Cognition {.tabset}

### PACC

```{r}
regression_analysis(data, 'NP_PACC_PACC96', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### FCTOTAL96

```{r}
regression_analysis(data, 'FCTOTAL96' , 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### MMSE

```{r}
regression_analysis(data, 'NP_MMSE_MMSE', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Digit

```{r}
regression_analysis(data, 'NP_DigitSym_DigitSym', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### LDL

```{r}
regression_analysis(data, 'NP_LogicMem_LM1ADR', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### YearsEd

```{r}
regression_analysis(data, 'YrsEd', 'delta', c('SEX', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

## Blood Biomarkers {.tabset}

### pTau217

```{r}
regression_analysis(data, 'p_tau217', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))

regression_analysis(data, 'p_tau217', 'delta', c('nfl', 'gfap'))
```

```{r}
fig <- scatter_plot <- ggplot(data, aes(x = delta, y = p_tau217)) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(x = 'Delta (years)',
         y = 'pTau217 (pg/mL)')

# Save 
ggsave("~/Downloads/pTau217_habs_cor.png", plot = fig, width = 4, height = 2.5, units = "in")
```

### pTau217 ratio

```{r}
regression_analysis(data, 'p_tau217_ratio', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))

regression_analysis(data, 'p_tau217_ratio', 'delta', c('nfl', 'gfap'))
```



### NFL

```{r}
regression_analysis(data, 'nfl', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))

regression_analysis(data, 'nfl', 'delta', c('p_tau217', 'gfap'))

```

### GFAP

```{r}
regression_analysis(data, 'gfap', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))

regression_analysis(data, 'gfap', 'delta', c('p_tau217', 'nfl'))
```


## PET Amyloid

### Composite

```{r}
regression_analysis(data, 'PIB_FS_SUVR_FLR', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

```{r}
fig <- scatter_plot <- ggplot(data, aes(x = delta, y = PIB_FS_SUVR_FLR)) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(x = 'Delta (years)',
         y = 'AB-PET SUVR')

# Save 
ggsave("~/Downloads/ab_habs_cor.png", plot = fig, width = 4, height = 2.5, units = "in")
```

## PET Tau {.tabset}

### Composite
```{r}
regression_analysis(data, 'tau_composite', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```
```{r}
fig <- scatter_plot <- ggplot(data, aes(x = delta, y = tau_composite)) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(x = 'Delta (years)',
         y = 'Tau-PET SUVR')

# Save 
ggsave("~/Downloads/tau_habs_cor.png", plot = fig, width = 4, height = 2.5, units = "in")
```


### Inferior temporal

```{r}
regression_analysis(data, 'inferiortemporal_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Inferior parietal

```{r}
regression_analysis(data, 'inferiorparietal_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Middle temporal

```{r}
regression_analysis(data, 'middletemporal_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Entorhinal

```{r}
regression_analysis(data, 'entorhinal_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Parahippocampal

```{r}
regression_analysis(data, 'parahippocampal_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Fusiform

```{r}
regression_analysis(data, 'fusiform_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Amygdala

```{r}
regression_analysis(data, 'amygdala_bh', 'delta',  c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```
