---
title: "Brain Age Longitudinal PACC"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    self_contained: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('BrainAgeLongitudinalPACC.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/BrainAgeLongitudinalPACC', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr);library(nlme);library(sjPlot);library(gtsummary); library(flextable)


knitr::opts_chunk$set(echo = FALSE, message = FALSE)
theme_set(theme_bw())
```

# Brain Age Longitudinal PACC

We will explore how baseline Brain Age delta moderates longitudinal PACC through a linear mixed effects model.

## A4 Structural BrainAge

```{r}

# Load data
data_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/longitudinal_PACC.csv')

# Load prs
prs_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/prs_0.5.csv')
data_A4 <- merge(data_A4, prs_A4, by = 'BID', all.x = TRUE)

# Load factors
factors_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/factors.csv')
#Extract only BID, pTau217, Amyloid Composite and Tau Composite
factors_A4 <- factors_A4 %>% 
  select(BID, pTau217, amyloid_composite, tau_composite)
data_A4 <- merge(data_A4, factors_A4, by = 'BID', all.x = TRUE)

# Load covariates
covars_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/covariates.csv')
covars_A4 <- covars_A4 %>% 
  select(BID, Amyloid)
data_A4 <- merge(data_A4, covars_A4, by = 'BID', all.x = TRUE)

# Filter subjects with less than 2 data points
data_A4 <- data_A4 %>% 
  arrange(BID, pacc_time) %>%  
  group_by(BID) %>% 
  add_tally(name = 'nTimePoints', wt = !is.na(PACC)) %>%
  filter(nTimePoints >= 2)

# Load deltas
delta_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/A4/structural_brainage/ageml/model_age/predicted_age_sex.csv')
delta_pet_A4 <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/A4/pet_brainage/ageml/model_age/predicted_age.csv')
delta_pet_A4 <- delta_pet_A4 %>% 
  select(BID, delta_all) %>% 
  dplyr::rename(delta_pet = delta_all)

# Merge the two datasets
data_A4 <- merge(data_A4, delta_A4, by = 'BID')
data_A4 <- merge(data_A4, delta_pet_A4, by = 'BID', all.x = TRUE)

# Arrange by pacc_time and calculate PACC_change_bl
data_A4 <- data_A4 %>% 
  arrange(BID, pacc_time) %>%
  group_by(BID) %>%
  mutate(PACC_bl = PACC[1]) %>%
  mutate(PACC_change_bl = PACC - PACC_bl)

# Create new variable delta_group where is 0 below younger and older above zero
data_A4 <- data_A4 %>% 
  mutate(delta_group = ifelse(delta_all < 0, 'younger', 'older')) %>%
  mutate(delta_pet_group = ifelse(delta_pet < 0, 'younger', 'older'))

# Create cohort variable
data_A4 <- data_A4 %>%
  group_by(BID) %>%
  dplyr::mutate(Cohort = case_when(
    SUBSTUDY == "SF" | SUBSTUDY == "LEARN" ~ "LEARN/SF",
    SUBSTUDY == "A4" & TX == "Placebo" ~ "A4 Placebo",
    SUBSTUDY == "A4" & TX == "Solanezumab" ~ "A4 Treated",
    TRUE ~ NA))

# Ensure variables are factors
data_A4 <- data_A4 %>%
  mutate_at(vars(SUBSTUDY, TX, Cohort, SEX, RACE, ETHNIC, APOE, APOEGN, QSVERSION_PACC), factor)

# Placebo data
data_placebo <- data_A4 %>% 
  filter(TX == 'Placebo' | SUBSTUDY != 'A4')
```

Plot PACC trajectories with spaghetti plot

```{r}
ggplot(data_placebo, aes(x = pacc_time, y = PACC, group = BID, colour = delta_group)) +
  geom_line() +
  geom_point() +
  labs(title = 'PACC Trajectories', x = 'Time from Baseline', y = 'PACC') +
  theme_minimal()
```

```{r}
ggplot(data_placebo, aes(x = pacc_time, y = PACC_change_bl, group = BID, colour = delta_group)) +
  geom_line() +
  geom_point() +
  labs(title = 'PACC Trajectories', x = 'Time from Baseline', y = 'PACC change from Basline') +
  theme_minimal()
```

Simplest linear mixed effects models with delta at baseline

PACC \~ pacc_time\^2 + pacc_timeXdelta + Cumulative_Dose_Scaled + QSVERSION_PACC

```{r}
lmm_results <- lme(PACC ~ I(pacc_time^2) + pacc_time*delta_all + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort, data = data_A4, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
p <- p +
  ggtitle(NULL) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/a4_pacc.png', plot = p, width = 3.5, height = 2)
```

### Interactions {.tabset}

#### pTau217

```{r}
lmm_results <- lme(PACC ~ I(pacc_time^2) + pacc_time*delta_all*pTau217 + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort, data = data_A4, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]', 'pTau217 [0.14, 0.28, 0.42]'))
levels(p$data$facet) <- c("pTau217 = 0.14", "pTau217 = 0.28", "pTau217 = 0.42")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/pTau217_a4_pacc.png', plot = p, width = 6, height = 3)
```

#### PET Amyloid

```{r}
lmm_results <- lme(PACC ~ I(pacc_time^2) + pacc_time*delta_all*amyloid_composite + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort, data = data_A4, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]', 'amyloid_composite [1.0, 1.17, 1.39]'))
levels(p$data$facet) <- c("AB-PET SUVR = 1.0", "AB-PET SUVR = 1.17", "AB-PET SUVR = 1.39")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/ab_a4_pacc.png', plot = p, width = 6, height = 3)
```

#### PET Tau

```{r}
lmm_results <- lme(PACC ~ I(pacc_time^2) + pacc_time*delta_all*tau_composite + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort, data = data_A4, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]', 'tau_composite [1.19, 1.4, 1.62]'))
levels(p$data$facet) <- c("Tau-PET SUVR = 1.19", "Tau-PET SUVR = 1.4", "Tau-PET SUVR = 1.62")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/tau_a4_pacc.png', plot = p, width = 6, height = 3)
```

Linear mixed effects models with delta at baseline

For our base model we will only use Learn and Placebo data and include as fixed effects the Cumulative Dose for those who opted into the treatmet at the end of the study and story version of the PACC as they are known to have an effect.

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*delta_all + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC , data = data_placebo, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'PACC_bl [-3, 0, 2]'), show.values = TRUE)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]', 'PACC_bl [-3, 0, 2]'), show.values = TRUE)
```

### Sensitivity Analysis

Linear mixed efffects models without splines to show worse fit.

Linear mixed effects models with change from baseline to show seperate effects of pacc at basline

Linear mixed effects model with splines

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_all + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC, data = data_placebo, random = ~pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
```

Linear mixed effects models with spline and Sex, Education Years, Basline Age, Apoe Status

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_all + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC + SEX*pacc_time + EDCCNTU*pacc_time + APOE*pacc_time + bl_pacc_age*pacc_time , data = data_placebo, random = ~pacc_time|BID, na.action = 'na.omit',control = lmeControl(opt = 'optim', msMaxIter = 100, msMaxEval = 100), method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
```

Linear mixed effects models with spline and all groups

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_all + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort, data = data_A4, random = ~pacc_time|BID, na.action = 'na.omit', method = c('ML'), control = lmeControl(opt = 'optim', msMaxIter = 100, msMaxEval = 100));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'Cohort'))
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
```

Linear mixed effects models with spline and all groups and interactions

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_all + pacc_time*PACC_bl + pacc_time*Cohort + Cumulative_Dose_Scaled + QSVERSION_PACC + SEX*pacc_time + EDCCNTU*pacc_time + APOE*pacc_time + bl_pacc_age*pacc_time, data = data_A4, random = ~pacc_time|BID, na.action = 'na.omit',control = lmeControl(opt = 'optim', msMaxIter = 100, msMaxEval = 100), method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_all [-3, 0, 3]'))
```

## A4 PET BrainAge

Plot PACC trajectories with spaghetti plot

```{r}
ggplot(subset(data_placebo, !is.na(delta_pet_group)), aes(x = pacc_time, y = PACC, group = BID, colour = delta_pet_group)) +
  geom_line() +
  geom_point() +
  labs(title = 'PACC Trajectories', x = 'Time from Baseline', y = 'PACC') +
  theme_minimal()
```

```{r}
ggplot(subset(data_placebo, !is.na(delta_pet_group)), aes(x = pacc_time, y = PACC_change_bl, group = BID, colour = delta_pet_group)) +
  geom_line() +
  geom_point() +
  labs(title = 'PACC Trajectories', x = 'Time from Baseline', y = 'PACC change from Basline') +
  theme_minimal()
```

Linear mixed effects models with delta at baseline

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*delta_pet + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC , data = data_placebo, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_pet [-3, 0, 3]'))
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'PACC_bl [-3, 0, 2]'), show.values = TRUE)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_pet [-3, 0, 3]', 'PACC_bl [-3, 0, 2]'), show.values = TRUE)
```

### Sensitivity Analysis

Linear mixed effects model with splines

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_pet + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC, data = data_placebo, random = ~pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_pet [-3, 0, 3]'))
```

Linear mixed effects models with spline and Sex, Education Years, Basline Age, Apoe Status

```{r}
lmm_results <- lme(PACC_change_bl ~ I(pacc_time^2) + pacc_time*delta_pet + pacc_time*PACC_bl + Cumulative_Dose_Scaled + QSVERSION_PACC + SEX*pacc_time + EDCCNTU*pacc_time + APOE*pacc_time + bl_pacc_age*pacc_time , data = data_placebo, random = ~pacc_time|BID, na.action = 'na.omit',control = lmeControl(opt = 'optim', msMaxIter = 100, msMaxEval = 100), method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'delta_pet [-3, 0, 3]'))
```

## A4 PRS BrainAge {.tabset}

Different linear mixed effect models exploring the interaciton of baseline age and PRS of different BrainAge models

### GM

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*GM_ZSCORE*bl_pacc_age + Cumulative_Dose_Scaled + QSVERSION_PACC , data = data_placebo, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
# Show if the interaction is significant
coef_table <- summary(lmm_results)$tTable
coef_table
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'bl_pacc_age', 'GM_ZSCORE [-1, 0, 1]'))
```

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*GM_ZSCORE*pTau217 + Cumulative_Dose_Scaled + QSVERSION_PACC + pacc_time*Cohort , data = data_A4, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
plot_model(lmm_results, type = c('pred'), terms = c('pacc_time', 'pTau217', 'GM_ZSCORE [-1, 0, 1]'))
```

### WM

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*WM_ZSCORE*bl_pacc_age + Cumulative_Dose_Scaled + QSVERSION_PACC , data = data_placebo, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
# Show if the interaction is significant
coef_table <- summary(lmm_results)$tTable
coef_table
```

### FC

```{r}
lmm_results <- lme(PACC_change_bl ~ pacc_time*FC_ZSCORE*bl_pacc_age + Cumulative_Dose_Scaled + QSVERSION_PACC , data = data_placebo, random = ~ 0 + pacc_time|BID, na.action = 'na.omit', method = c('ML'));
# Show if the interaction is significant
coef_table <- summary(lmm_results)$tTable
coef_table
```

## HABS

```{r}

# Load data
data_habs <- read.csv('~/Documents/Projects/AgingBrainStudy/data/HABS/raw/longitudinal_PACC.csv')

# Convert NP_SessionDate to Date
data_habs$NP_SessionDate <- as.Date(data_habs$NP_SessionDate)

# Filter subjects with less than 2 data points
data_habs <- data_habs %>% 
  arrange(SubjIDshort, NP_SessionDate) %>%  
  group_by(SubjIDshort) %>% 
  add_tally(name = 'nTimePoints', wt = !is.na(NP_PACC_PACC5)) %>%
  filter(nTimePoints >= 2)

# Calculate time form base line
data_habs <- data_habs %>% 
  group_by(SubjIDshort) %>% 
  mutate(time_from_bl = as.numeric(NP_SessionDate - min(NP_SessionDate, na.rm = TRUE))/365.25)

# Create baseline data set
baseline_habs <- data_habs %>% 
  group_by(SubjIDshort) %>%
  arrange(NP_SessionDate) %>%
  slice(1) %>% #extract first rows - can also use slice_head() for first and slice_tail() for last
  rename(BL_Age = NP_Age) %>%
  rename(BL_PACC = NP_PACC_PACC5) %>%
  ungroup()

# Add age at baseline of subjects
data_habs <- data_habs %>%
  left_join(baseline_habs %>% select(SubjIDshort, BL_Age, BL_PACC), by = 'SubjIDshort')

# Load csv file 
deltas_habs <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/HABS/structural_brainage/ageml/model_age/predicted_age.csv')

# Merge the two datasets
data_habs <- merge(data_habs, deltas_habs, by = 'SubjIDshort', all.x = TRUE)

# Create new variable delta_group where is 0 below younger and older above zero
data_habs <- data_habs %>% 
  mutate(delta_group = ifelse(delta_all < 0, 'younger', 'older'))

# Add factors
factors_habs <- read.csv('~/Documents/Projects/AgingBrainStudy/data/HABS/processed/factors.csv')
# Only pTau217, amyloid composite and tau composite
factors_habs <- factors_habs %>% 
  select(SubjIDshort, ptau, p_tau217_ratio, PIB_FS_SUVR_FLR, PIB_FS_SUVR_Group, tau_composite)
data_habs <- merge(data_habs, factors_habs, by = 'SubjIDshort', all.x = TRUE)
```

Plot PACC trajectories with spaghetti plot

```{r}
ggplot(subset(data_habs, !is.na(delta_group)), aes(x = time_from_bl, y = NP_PACC_PACC5, group = SubjIDshort, colour = delta_group)) +
  geom_line() +
  geom_point() +
  labs(title = 'PACC Trajectories', x = 'Time from Baseline', y = 'PACC') +
  theme_minimal()
```

Linear mixed effects models with delta at baseline

```{r}
lmm_results <- lme(NP_PACC_PACC5 ~ time_from_bl*delta_all, data = data_habs, random = ~time_from_bl|SubjIDshort, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('time_from_bl', 'delta_all [-4, 0, 4]'))
p <- p +
  ggtitle(NULL) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/habs_pacc.png', plot = p, width = 3.5, height = 2)
```

### Interactions {.tabset}

#### pTau217

```{r}
lmm_results <- lme(NP_PACC_PACC5 ~ time_from_bl*delta_all*p_tau217_ratio, data = data_habs, random = ~time_from_bl|SubjIDshort, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('time_from_bl', 'delta_all [-4, 0, 4]', 'p_tau217_ratio [0.71, 3.46, 6.21]'))
levels(p$data$facet) <- c("pTau217 = 0.71", "pTau217 = 3.46", "pTau217 = 6.21")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  ggtitle(NULL) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/pTau217_habs_pacc.png', plot = p, width = 6, height = 3)
```

#### PET Amyloid

```{r}
lmm_results <- lme(NP_PACC_PACC5 ~ time_from_bl*delta_all*PIB_FS_SUVR_FLR, data = data_habs, random = ~time_from_bl|SubjIDshort, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('time_from_bl', 'delta_all [-4, 0, 4]', 'PIB_FS_SUVR_FLR [1.0, 1.24, 1.49]'))
levels(p$data$facet) <- c("AB-PET SUVR = 1.0", "AB-PET SUVR = 1.24", "AB-PET SUVR = 1.49")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/ab_habs_pacc.png', plot = p, width = 6, height = 3)
```

#### PET Tau

```{r}
lmm_results <- lme(NP_PACC_PACC5 ~ time_from_bl*delta_all*tau_composite, data = data_habs, random = ~time_from_bl|SubjIDshort, na.action = 'na.omit', method = c('ML'));
summary(lmm_results)
p <- plot_model(lmm_results, type = c('pred'), terms = c('time_from_bl', 'delta_all [-4, 0, 4]', 'tau_composite [1.17, 1.35, 1.52]'))
levels(p$data$facet) <- c("Tau-PET SUVR = 1.17", "Tau-PET SUVR = 1.35", "Tau-PET SUVR = 1.52")
p <- p +
  ggtitle(NULL) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(color = 'Delta (years)',
      x = 'Time from baseline (years)',
      y = 'PACC')
p
ggsave('~/Downloads/tau_habs_pacc.png', plot = p, width = 6, height = 3)
```

## Demographics

```{r}
data_A4_demo <- data_A4 %>% 
  arrange(BID, pacc_time) %>%  
  group_by(BID) %>%
  mutate(follow_up_time = max(pacc_time, na.rm = TRUE)) %>%
  slice(1) %>% #extract first rows - can also use slice_head() for first and slice_tail() for last
  select(BID, AGEYR, SEX, EDCCNTU, APOE, Amyloid, follow_up_time) %>%
  # Add a column Cohort that has in its entry A4/Learn
  mutate(Cohort = 'A4/Learn') %>%
  mutate(Amyloid = ifelse(Amyloid == 1.0, 'positive', 'negative'))

data_HABS_demo <- data_habs %>% 
  arrange(SubjIDshort, NP_SessionDate) %>%  
  group_by(SubjIDshort) %>% 
  mutate(follow_up_time = max(time_from_bl, na.rm = TRUE)) %>%
  slice(1) %>% #extract first rows - can also use slice_head() for first and slice_tail() for last
  select(SubjIDshort, NP_Age, SEX, YrsEd, E4_Status, PIB_FS_SUVR_Group, follow_up_time) %>%
  rename(BID=SubjIDshort, AGEYR = NP_Age, EDCCNTU = YrsEd, APOE = E4_Status, Amyloid = PIB_FS_SUVR_Group) %>%
  mutate(Cohort = 'HABS') %>%
  # Convert PIB- and PIB + to 0 and 1
  mutate(Amyloid = ifelse(Amyloid == 'PIB-', 'negative', 'positive')) %>%
  mutate(SEX = case_when(
    SEX == "F" ~ "Female",
    SEX == "M" ~ "Male"))

habs_t <- tbl_summary(data_HABS_demo,
            include = c(AGEYR, SEX, EDCCNTU, APOE, Amyloid, follow_up_time),
            missing = "no",
            statistic = list(all_continuous()  ~ c("{mean} ({sd})")),
            label = list(AGEYR ~ "Age", EDCCNTU ~ "Education (Years)", SEX ~ "Sex", follow_up_time ~ "Follow-up Time (Years)"),
            digits = list(all_continuous() ~ 2, all_categorical() ~ c(0,2)),
            by = Amyloid) %>%
  bold_labels()

A4_t <- tbl_summary(data_A4_demo,
            include = c(AGEYR, SEX, EDCCNTU, APOE, Amyloid, follow_up_time),
            missing = "no",
            statistic = list(all_continuous()  ~ c("{mean} ({sd})")),
            label = list(AGEYR ~ "Age", EDCCNTU ~ "Education (Years)", SEX ~ "Sex", follow_up_time ~ "Follow-up Time (Years)"),
            digits = list(all_continuous() ~ 2, all_categorical() ~ c(0,2)),
            by = Amyloid) %>%
  bold_labels()

table <- tbl_merge(list(A4_t, habs_t), tab_spanner = c("A4/LEARN", "HABS")) 

table

table %>%
  as_flex_table() %>% 
  fontsize(size = 14, part = "all") %>%
  line_spacing(space = 1.5, part = "all") %>%
  autofit() %>%
  save_as_docx(path = "~/Downloads/DemographicsTable.docx")
  
```
