---
title: "HABSPETBrainAge"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('HABSPETBrainAge.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/HABSPETBrainAge', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr); library(sjPlot)


knitr::opts_chunk$set(echo = FALSE, message = FALSE)
theme_set(theme_bw())
```

# BrainAge modelling in HABS

A linear regression BrainAge model was trained on e4 non-carriers. The model was then applied to e4 carriers.

The model was trained using the following regions:

-   Amygdala 
-   Inferior parietal
-   Inferior temporal
-   Middle temporal
-   Entorhinal
-   Parahippocampal
-   Fusiform

Command use: model_age -o analysis/HABS/pet_brainage -f data/HABS/processed/pet_features.csv --clinical data/HABS/processed/e4.csv

```{r, include = FALSE}
# Load deltas
deltas <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/HABS/pet_brainage/ageml/model_age/predicted_age.csv')
deltas <- deltas %>% 
  rename(delta = delta_all, predicted_age = predicted_age_all)
delta_struct <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/HABS/structural_brainage/ageml/model_age/predicted_age.csv')
names(delta_struct)[names(delta_struct) == 'delta_all'] <- 'delta_struct'
data <- merge(deltas, delta_struct[, c('SubjIDshort', 'delta_struct')], by = 'SubjIDshort', all.x = TRUE)

# Obtain deltas trained with A4
deltas_trained_wA4 <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/A4_HABS_joint/pet_brainage/ageml/model_age/predicted_age.csv')
deltas_trained_wA4 <- deltas_trained_wA4[deltas_trained_wA4$SubjIDshort %in% grep('_', deltas_trained_wA4$SubjIDshort, value = TRUE),]
deltas_trained_wA4 <- deltas_trained_wA4 %>% 
  rename(predicted_age_trained_wA4 = predicted_age_all, delta_wA4 = delta_all) %>%
  select(SubjIDshort, predicted_age_trained_wA4, delta_wA4)
data <- merge(data, deltas_trained_wA4, by = 'SubjIDshort', all = TRUE)

# Load factors
factors <- read.csv('~/Documents/Projects/AgingBrainStudy/data/HABS/processed/factors.csv', na.strings = c("", " ", "NaN", "NA", NA))
data <- merge(data, factors, by = 'SubjIDshort', all = TRUE)
```

## Comparison of BrainAge models

We also trained a model on A4 and applied it to HABS.

Command used: model_age -o analysis/A4_HABS_joint/pet_brainage -f data/A4_HABS_joint/processed/pet_features.csv --clinical data/A4_HABS_joint/A4_as_cn.csv

```{r}
# Plot deltas vs deltas_trained_wA4
ggplot(data = data, aes(x = delta, y = delta_wA4)) +
  geom_point() +
  stat_cor() +
  labs(x = 'Delta', y = 'Delta A4 trained', title = 'Comparison of BrainAge models')

# histogram of deltas coloured by model
ggplot(data = data, aes(x = delta, fill = 'delta')) +
  geom_histogram(alpha = 0.5, position = 'identity') +
  geom_histogram(data = data, aes(x = delta_wA4, fill = 'delta_wA4'), alpha = 0.5, position = 'identity') +
  labs(x = 'Delta', y = 'Count', title = 'Histogram of deltas coloured by model') +
  scale_fill_manual(values = c('blue', 'red'))
```

```{r}
# Caculate mean absolute error for each model
mae <- data %>% 
  summarise(mae = mean(abs(age- predicted_age)),
            mae_trained_wA4 = mean(abs(age - predicted_age_trained_wA4), na.rm = TRUE))
mae
std <- data %>% 
  summarise(std = sd(delta),
            std_trained_wA4 = sd(delta_wA4, na.rm = TRUE))
std
```

## PET BrainAge vs Structural BrainAge

```{r}
ggplot(data, aes(x = delta, y = delta_struct)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  stat_cor(method = 'pearson') +
  labs(title = 'PET BrainAge vs Structural BrainAge',
       x = 'PET BrainAge',
       y = 'Structural BrainAge')
```


```{r}
regression_analysis <- function(data, x, y, covars) {
  
  # Filter nan values
  data <- data[complete.cases(data[, c(x, y, covars)]), ]
  
  # Scatter plot with correlation
  scatter_plot <- ggplot(data, aes(x = !!sym(y), y = !!sym(x))) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
    
  # Create a formula for the regression model
  covar_formula <- paste(covars, collapse = " + ")
  full_formula <- as.formula(paste(x, "~", y, "+", covar_formula))
  
  # Linear regression model
  ols_results <- lm(as.formula(paste(x, "~", y)), data = data)
  ols_full_results <- lm(full_formula, data = data)
  
  # Model summary
  model_summary <- summary(ols_results)
  model_full_summary <- summary(ols_full_results)
  
  # Plot model
  pred_plot <- plot_model(ols_results, type = c('pred'), terms = c(y), show.values = TRUE) +
    geom_point(data = data, aes(x = !!sym(y), y = !!sym(x)), 
             inherit.aes = F) + 
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
  
  # Return a list with all results
  return(list(
    scatter_plot = scatter_plot,
    simple_model = ols_results,
    regression_model = ols_full_results,
    model_summary = model_summary,
    model_full_summary = model_full_summary,
    predicted_values_plot = pred_plot
  ))
}
```

## Delta differences {.tabset}

### E4 status

```{r}
ggplot(subset(data, !is.na(E4_Status)), aes(x = E4_Status, y = delta)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by E4 status',
       x = 'E4 status',
       y = 'Delta')
```

### Amyloid status

```{r}
ggplot(subset(data, !is.na(PIB_FS_SUVR_Group)), aes(x = PIB_FS_SUVR_Group, y = delta)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by Amyloid status',
       x = 'Amyloid status',
       y = 'Delta')
```

## Cognition {.tabset}

### PACC

```{r}
regression_analysis(data, 'NP_PACC_PACC96', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### FCTOTAL96

```{r}
regression_analysis(data, 'FCTOTAL96' , 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### MMSE

```{r}
regression_analysis(data, 'NP_MMSE_MMSE', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### Digit

```{r}
regression_analysis(data, 'NP_DigitSym_DigitSym', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```

### LDL

```{r}
regression_analysis(data, 'NP_LogicMem_LM1ADR', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```


## PET Amyloid

### Composite

```{r}
regression_analysis(data, 'PIB_FS_SUVR_FLR', 'delta', c('SEX', 'YrsEd', 'E4_Status', 'PIB_FS_SUVR_Group', 'age'))
```
