---
title: "A4 Pet BrainAge"
author: "Jorge Garcia Condado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render('A4PETBrainAge.Rmd',
      output_file = paste('~/Documents/Projects/AgingBrainStudy/results/PETA4BrainAge', '.html', sep=''))
      })
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2); library(ggpubr); library(dplyr);library(sjPlot)


knitr::opts_chunk$set(echo = FALSE, message = FALSE)
theme_set(theme_bw())
```

# PET BrainAge modelling in A4/Elearn study

A linear regression BrainAge model was trained on e4 non-carriers. The model was then applied to e4 carriers.

The model was trained using the following regions:

-   Amygdala 
-   Inferior parietal
-   Inferior temporal
-   Middle temporal
-   Entorhinal
-   Parahippocampal
-   Fusiform

Command use: model_age -o analysis/A4/pet_brainage -f data/A4/processed/pet_features.csv --clinical data/A4/processed/e4.csv

```{r, include = FALSE}

# Load data
deltas <- read.csv('~/Documents/Projects/AgingbrainStudy/analysis/A4/pet_brainage/ageml/model_age/predicted_age.csv')
delta_struct <- read.csv('~/Documents/Projects/AgingBrainStudy/analysis/A4/structural_brainage/ageml/model_age/predicted_age_sex.csv')
names(delta_struct)[names(delta_struct) == 'delta_all'] <- 'delta_struct_all'
data <- merge(deltas, delta_struct, by = 'BID', all.x = TRUE)

# Load factors
factors <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/factors.csv', na.strings = c("", " ", "NaN", "NA", NA))
data <- merge(data, factors, by = 'BID', all.x = TRUE)

# Amyloid status
covars <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/processed/covariates.csv', na.strings = c("", " ", "NaN", "NA", NA))
covars$Amyloid <- ifelse(covars$Amyloid == 1, 'positive', 'negative')
covars$e4 <- ifelse(covars$e4 == 1, 'carrier', 'non-carrier')
data <- merge(data, covars[, c('BID', 'Amyloid', 'e4')], by = 'BID', all.x = TRUE)

# Demographics
demographics <- read.csv('~/Documents/Projects/AgingBrainStudy/data/A4/raw/SUBJINFO.csv', na.strings = c("", " ", "NaN", "NA", NA))
demographics <- demographics %>%
  rename(YearsEd = EDCCNTU) %>%
  rename(BL_Age = AGEYR)
data <- merge(data, demographics, by = 'BID', all.x = TRUE)

```

```{r}
regression_analysis <- function(data, x, y, covars) {
  
  # Filter nan values
  data <- data[complete.cases(data[, c(x, y, covars)]), ]
  
  # Scatter plot with correlation
  scatter_plot <- ggplot(data, aes(x = !!sym(y), y = !!sym(x))) +
    geom_point() +
    geom_smooth(method = 'lm', formula = y ~ x) +
    stat_cor(method = 'pearson') +
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
    
  # Create a formula for the regression model
  covar_formula <- paste(covars, collapse = " + ")
  full_formula <- as.formula(paste(x, "~", y, "+", covar_formula))
  
  # Linear regression model
  ols_results <- lm(as.formula(paste(x, "~", y)), data = data)
  ols_full_results <- lm(full_formula, data = data)
  
  # Model summary
  model_summary <- summary(ols_results)
  model_full_summary <- summary(ols_full_results)
  
  # Plot model
  pred_plot <- plot_model(ols_results, type = c('pred'), terms = c(y), show.values = TRUE) +
    geom_point(data = data, aes(x = !!sym(y), y = !!sym(x)), 
             inherit.aes = F) + 
    labs(title = paste(x, "vs", y),
         x = y,
         y = x)
  
  # Return a list with all results
  return(list(
    scatter_plot = scatter_plot,
    simple_model = ols_results,
    regression_model = ols_full_results,
    model_summary = model_summary,
    model_full_summary = model_full_summary,
    predicted_values_plot = pred_plot
  ))
}
```

## PET BrainAge vs Structural BrainAge

```{r}
ggplot(data, aes(x = delta_all, y = delta_struct_all)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  stat_cor(method = 'pearson') +
  labs(title = 'PET BrainAge vs Structural BrainAge',
       x = 'PET BrainAge',
       y = 'Structural BrainAge')
```

## Delta differences {.tabset}

### E4 status
```{r}
ggplot(subset(data, !is.na(e4)), aes(x = e4, y = delta_all)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by E4 status',
       x = 'E4 status',
       y = 'Delta')
```

### Amyloid status
```{r}
ggplot(subset(data, !is.na(Amyloid)), aes(x = Amyloid, y = delta_all)) +
  geom_boxplot() +
  stat_compare_means(method = 't.test') +
  labs(title = 'Delta differences by Amyloid status',
       x = 'Amyloid status',
       y = 'Delta')
```


## Cognition {.tabset}

### PACC

```{r}
regression_analysis(data, 'PACC', 'delta_all', c('SEX', 'YearsEd', 'e4', 'Amyloid', 'BL_Age'))
```

### FCTOTAL96

```{r}
regression_analysis(data, 'FCTOTAL96' , 'delta_all', c('SEX', 'YearsEd', 'e4', 'Amyloid', 'BL_Age'))
```

### MMSE

```{r}
regression_analysis(data, 'MMSCORE', 'delta_all', c('SEX', 'YearsEd', 'e4', 'Amyloid', 'BL_Age'))
```

### Digit

```{r}
regression_analysis(data, 'DIGITTOTAL', 'delta_all', c('SEX', 'YearsEd', 'e4', 'Amyloid', 'BL_Age'))
```

### LDL

```{r}
regression_analysis(data, 'LDELTOTAL', 'delta_all', c('SEX', 'YearsEd', 'e4', 'Amyloid', 'BL_Age'))
```

## Blood Biomarkers {.tabset}

### pTau217

```{r}
regression_analysis(data, 'pTau217', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))

regression_analysis(data, 'pTau217', 'delta_all', c('NF.L', 'AB42.AB40', 'GFAP'))

regression_analysis(data, 'pTau217', 'delta_all', c('NF.L', 'AB42.AB40', 'GFAP', 'TPP181'))
```

### tpp181

```{r}
regression_analysis(data, 'TPP181', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))

regression_analysis(data, 'TPP181', 'delta_all', c('NF.L', 'AB42.AB40', 'GFAP'))

regression_analysis(data, 'TPP181', 'delta_all', c('NF.L', 'AB42.AB40', 'GFAP', 'pTau217'))
```

### NFL

```{r}
regression_analysis(data, 'NF.L', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))

regression_analysis(data, 'NF.L', 'delta_all', c('pTau217', 'TPP181', 'AB42.AB40', 'GFAP'))

```

### GFAP

```{r}
regression_analysis(data, 'GFAP', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))

regression_analysis(data, 'GFAP', 'delta_all', c('pTau217', 'TPP181', 'AB42.AB40', 'NF.L'))
```

### AB42/40

```{r}
regression_analysis(data, 'AB42.AB40', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))

regression_analysis(data, 'AB42.AB40', 'delta_all', c('pTau217', 'TPP181', 'NF.L', 'GFAP'))
```

## PET Amyloid

### Composite

```{r}
regression_analysis(data, 'amyloid_composite', 'delta_all', c('SEX', 'YearsEd', 'e4', 'BL_Age'))
```

### Comparison to Tau Composite

```{r}
regression_analysis(data, 'amyloid_composite', 'tau_composite', c('SEX', 'YearsEd', 'e4', 'BL_Age'))
```

```{r}
regression_analysis(data, 'pTau217', 'tau_composite', c('SEX', 'YearsEd', 'e4', 'BL_Age'))
```